The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-11-15 script/en/chapters/paper_and_pencil.tex
CT script en: Sage example on the affine cipher

@@

>>>>

\item {\bf Affine cipher}: This is a generalization of the shift
   cipher. A plaintext character is first substituted for another
   character and then the result is encrypted using the shift
   cipher. The affine cipher is so named because its encryption and
   decryption functions are affine or linear functions.

@@

% ---------------------------------------------------------------------------
\newpage
\subsection{Affine cipher}

The affine cipher is implemented in the Sage class
\begin{center}
\verb!sage.crypto.classical.AffineCryptosystem!
\end{center}
In the following example, we construct an affine cipher with key
$(3, 13)$ and use this key to encrypt a plaintext. The plaintext is
then decrypted and the result is compared to the original plaintext.

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
sage: # create an affine cipher
sage: affineCipher = AffineCryptosystem(AlphabeticStrings()); affineCipher
Affine cryptosystem on Free alphabetic string monoid on A-Z
sage: plaintext = affineCipher.encoding("The affine cryptosystem.")
sage: plaintext
THEAFFINECRYPTOSYSTEM
sage: 
sage: # encrypt the plaintext using the key (3, 13)
sage: a, b = (3, 13)
sage: ciphertext = affineCipher.enciphering(a, b, plaintext)
sage: ciphertext
SIZNCCLAZTMHGSDPHPSZX
sage: 
sage: # decrypt the ciphertext and make sure that it is equivalent to the original plaintext
sage: P = affineCipher.deciphering(a, b, ciphertext)
sage: P
THEAFFINECRYPTOSYSTEM
sage: P == plaintext
True
\end{Verbatim}
\caption{An affine cipher with key $(3, 13)$.}
\end{sagecode}
